---
title: "Exercises: Vectors and functions"
author: "Shawn Santo"
date: ""
output: 
  html_document:
    css: "exercises.css"
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: false
    df_print: paged
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      comment = "#>", highlight = TRUE,
                      fig.align = "center")
```

# Exercise 1 {.tabset .tabset-fade .tabset-pills}

## Problem

What is the type of each vector below? Check your answer in R.

```{r eval=FALSE}
c(4L, 16, 0)
c(NaN, NA, -Inf)
c(NA, TRUE, FALSE, "TRUE")
c(pi, NaN, NA)
```

## Solution

```{r}
typeof(c(4L, 16, 0))
typeof(c(NaN, NA, -Inf))
typeof(c(NA, TRUE, FALSE, "TRUE"))
typeof(c(pi, NaN, NA))
```

# Exercise 2 {.tabset .tabset-fade .tabset-pills}

## Problem

Write a conditional statement that prints "Can't proceed NA or NaN present!"
if a vector contains `NA` or `NaN`. Test your code with vectors `x` and `y`
below.

```{r}
x <- NA
y <- c(1:5, NaN, NA, sqrt(3))
```

## Solution

```{r}
x <- NA
y <- c(1:5, NaN, NA, sqrt(3))
if (any(is.na(x))) {print("Can't proceed NA or NaN present!")}
if (any(is.na(y))) {print("Can't proceed NA or NaN present!")}
```

# Exercise 3 {.tabset .tabset-fade .tabset-pills}

## Problem

Create a list in R that has a similar structure as the JSON product-order data
below.

```
[
 {
  "id": {
    "oid": "5968dd23fc13ae04d9000001"
  },
  "product_name": "sildenafil citrate",
  "supplier": "Wisozk Inc",
  "quantity": 261,
  "unit_cost": "$10.47"
 },
 
 {
  "id": {
    "oid": "5968dd23fc13ae04d9000002"
  },
  "product_name": "Mountain Juniperus ashei",
  "supplier": "Keebler-Hilpert",
  "quantity": 292,
  "unit_cost": "$8.74"
 }
]
```

## Solution

```{r}
x <- list(
  list(
    id           = list(oid = "5968dd23fc13ae04d9000001"),
    product_name = "sildenafil citrate",
    supplier     = "Wisozk Inc",
    quantity     = 261,
    unit_cost    = "$10.47"
  ),
  list(
    id           = list(oid = "5968dd23fc13ae04d9000002"),
    product_name = "Mountain Juniperus ashei",
    supplier     = "Keebler-Hilpert",
    quantity     = 292,
    unit_cost    = "$8.74"
  )
)

str(x)
```

# Exercise 4 {.tabset .tabset-fade .tabset-pills}

## Problem

Write a function called `get_vector_properties()` that takes an atomic vector
and returns the following information in the format of a list: the vector's
type, length, and element names. *Hint*: element names can be retrieved with
function `names()`.

```{r echo=FALSE}
get_vector_properties <- function(x) {
  
  list(data_type = typeof(x),
       length    = length(x),
       names     = names(x)
       )
}
```

```{r}
get_vector_properties(c(5, 10, -5, 8, 11, 0))
get_vector_properties(c(a = "-1", b = "0", c = "1"))
```

## Solution

```{r}
get_vector_properties <- function(x) {
  
  list(data_type = typeof(x),
       length    = length(x),
       names     = names(x)
       )
}
```

# Exercise 5 {.tabset .tabset-fade .tabset-pills}

## Problem

We have seen the group membership operator `%in%`. Write an infix function
named `%!in%` that returns `TRUE` for each component in the left operand that
is not in the right operand. For example,

```{r echo=FALSE}
`%!in%` <- function(a, b) {
  !(a %in% b)
}
```

```{r}
x <- c(4, -4, 9, 33, 1)
y <- c(1, 44, 1, 0, 33, 10, 62)

x %!in% y
```

## Solution

```{r}
`%!in%` <- function(a, b) {
  !(a %in% b)
}
```

```{r}
x <- c(4, -4, 9, 33, 1)
y <- c(1, 44, 1, 0, 33, 10, 62)

x %!in% y
```

# Exercise 6 {.tabset .tabset-fade .tabset-pills}

*This function requires subsetting. If you can't get this to work, that's okay. 
We'll cover subsetting next week.*

## Problem

Write a function called `seq_fib()`. Function `seq_fib()` should take one 
argument, `n` - an atomic numeric vector of length 1, and return the first `n` 
[Fibonacci numbers](https://en.wikipedia.org/wiki/Fibonacci_number)
as an atomic vector. Some examples of the function in 
action are given below.

```{r echo=FALSE}
seq_fib <- function(n) {
  # check for n to be greater than or equal to 1
  stopifnot(length(n) == 1, n >= 1, n %/% 1 == n)
  
  fib.vec <- c(0, 1)
  if (n <= 2) { # n = 1 or n = 2
    fib.vec <- fib.vec[1:n]
  } else {
     for (i in 3:n) { # n >= 3
       fib.vec[i] <- fib.vec[i - 2] + fib.vec[i - 1]
     }
  }
  return(fib.vec)
}
```

#### Valid inputs

```{r results='hold'}
seq_fib(n = 1)
seq_fib(n = 2)
seq_fib(n = 7)
seq_fib(n = 10)
```

#### Invalid inputs

```{r error=TRUE, results='hold'}
seq_fib(n = -2)
seq_fib(n = c(2, 3, 10))
```

Did you remember to handle situations such as `seq_fib(n = 4.9)`?

## Solution

```{r}
seq_fib <- function(n) {
  # check for n to be greater than or equal to 1
  stopifnot(length(n) == 1, n >= 1, n %/% 1 == n)
  
  fib.vec <- c(0, 1)
  if (n <= 2) { # n = 1 or n = 2
    fib.vec <- fib.vec[1:n]
  } else {
     for (i in 3:n) { # n >= 3
       fib.vec[i] <- fib.vec[i - 2] + fib.vec[i - 1]
     }
  }
  return(fib.vec)
}
```

