---
title: "Exercises: Web APIs"
author: "Shawn Santo"
date: ""
output: 
  html_document:
    css: "exercises.css"
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: false
    df_print: paged
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      comment = "#>", highlight = TRUE,
                      fig.align = "center")
```

# Packages

```{r}
library(jsonlite)
library(tidyverse)
```

# Data

Use the [Open Brewery API](https://www.openbrewerydb.org/) to answer the 
following questions. You'll need to read through their documentation to 
get an idea for how it is structured.

# Exercise 1 {.tabset .tabset-fade .tabset-pills}

## Problem

How many breweries are located in Durham, NC?

## Solution

```{r}
base_url <- "https://api.openbrewerydb.org/breweries"

# build query part
query <- "?by_state=north+carolina&by_city=durham&per_page=50"

read_json(str_c(base_url, query)) %>% 
  length()
```

# Exercise 2 {.tabset .tabset-fade .tabset-pills}

## Problem

Which city in North Carolina has the most micro breweries? How many micro
breweries do they have?

## Solution

We'll use the `base_url` object created in the solution to Exercise 1.

```{r}
get_nc_brew <- function(base, page) {
  query <- str_c("?by_state=north_carolina&by_type=micro&page=", 
                 page, "&per_page=50")
  fromJSON(str_c(base_url, query))
}

# iterate over 10 pages, since we don't know how many pages there are
nc_micro <- map(1:10, get_nc_brew, base = base_url) %>% 
  map_df(rbind) %>% 
  as_tibble()

# determine which city has the most micro breweries
nc_micro %>% 
  group_by(city) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count)) %>% 
  slice(1)
```

# Exercise 3 {.tabset .tabset-fade .tabset-pills}

## Problem

In what cities are Founders, Yuengling, and Boulevard brewed?

## Solution

We'll use the `base_url` object created in the solution to Exercise 1.

```{r}
brew <- c("founders", "yuengling", "boulevard")

get_city_brew <- function(co) {
  query <- str_c("?by_name=", co)
  fromJSON(str_c(base_url, query)) %>% 
  pull(city) %>% 
  unique()
}

map(brew, get_city_brew) %>% 
  set_names(brew)
```

# Exercise 4 {.tabset .tabset-fade .tabset-pills}

## Problem

Which state has the most breweries?

## Solution

We'll use the `base_url` object created in the solution to Exercise 1.

```{r}
get_brew <- function(page) {
  query <- str_c("?page=", page, "&per_page=50")
  fromJSON(str_c(base_url, query)) %>% 
    as_tibble()
}

map_df(1:200, get_brew) %>% 
  count(state) %>% 
  arrange(desc(n)) %>% 
  slice(1)
```
