---
title: "Exercises: String manipulation and regexes"
author: "Shawn Santo"
date: ""
output: 
  html_document:
    css: "exercises.css"
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: false
    df_print: paged
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      comment = "#>", highlight = TRUE,
                      fig.align = "center")
```

# Packages

```{r}
library(tidyverse)
```

# Exercise 1 {.tabset .tabset-fade .tabset-pills}

## Problem

1. What `pattern` do you need to match a literal `\` in the string below?
    ```{r eval=FALSE}
    str_view_all("find the \\ in this string", pattern = )
    ```
2. Can you extract the arithmetic operations in this string?
    ```{r eval=FALSE}
    str_extract_all("4 + 3 ^ 2 - 10", pattern = )
    ```
3. What `pattern` do you need to match every character and `?` combination?
    ```{r eval=FALSE}
    str_view_all("to get help? in r you can use a ?, maybe this
                 works in other languages?", pattern = )
    ```

## Solution

```{r}
str_view_all("find the \\ in this string", pattern = "\\\\")

str_extract_all("4 + 3 ^ 2 - 10", pattern = "[\\+\\^\\-]")

str_view_all("to get help? in r you can use a ?, maybe this
                 works in other languages?", pattern = ".\\?")
```


# Exercise 2 {.tabset .tabset-fade .tabset-pills}

## Problem

Write a regular expression to match a

1. social security number of the form ###-##-####,

2. phone number of the form (###) ###-####,

3. license plate of the form AAA ####.

Test your regexs on some examples with `str_detect()` or `str_view()`.

## Solution

```{r}
x <- "My info is as follows. Cell: (432)-431-1512. Social security: 432-11-1990"
y <- "Vehicle info: AEF 2348"

# not the most efficient
str_view_all(x, "[0-9][0-9][0-9]-[0-9][0-9]-[0-9][0-9][0-9][0-9]")

str_view_all(x, "\\([0-9][0-9][0-9]\\)-[0-9][0-9][0-9]-[0-9][0-9][0-9][0-9]")

str_view_all(y, "[A-Z][A-Z][A-Z] [0-9][0-9][0-9][0-9]")
```

# Exercise 3 {.tabset .tabset-fade .tabset-pills}

## Problem

1. Write a regular expression that will extract all phone numbers contained 
   in the vector above.

2. Once that works use groups to extracts the area code separately from the rest
   of the phone number.

```{r}
text <- c(
  "apple", 
  "219 733 8965", 
  "329-293-8753",
  "Work: (579) 499-7527; Home: (543) 355 3679"
)
```

## Solution

```{r}
str_extract_all(text, "\\(?\\d{3}\\)?[ -]\\d{3}[ -]\\d{4}")

str_match_all(text, "\\(?(\\d{3})\\)?[ -]\\d{3}[ -]\\d{4}") %>% 
  purrr::map(~ .[, 2]) %>% 
  unlist()
```

# Exercise 4 {.tabset .tabset-fade .tabset-pills}

## Problem

Extract every word "fruit" or "flies" from `phrases`.

```{r}
phrases <-  c("time flies when you're having fun in 521",
              "fruit flies when you throw it",
              "a fruit fly is a beautiful creature",
              "how do you spell fruitfly?")
```

## Solution

```{r}
phrases %>% 
  str_extract_all(pattern = "fruit|flies")
```

# Exercise 5 {.tabset .tabset-fade .tabset-pills}

## Problem

Extract the numeric values from the tongue twister below.

```{r}
twister <- paste("Something in a 30 acre thermal thicket of thorns and",
                 "thistles thumped and thundered threatening the 3-D",
                 "thoughts of Matthew the thug - although, theatrically,",
                 "it was only the 13000 thistles and thorns through the",
                 "underneath of his thigh that the 30 year old thug",
                 "thought of that morning.", sep = " ")
```

## Solution

```{r}
twister %>% 
  str_extract_all(pattern = "[:digit:]+")
```