---
title: "Exercises: Web scraping part I"
author: "Shawn Santo"
date: ""
output: 
  html_document:
    css: "exercises.css"
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: false
    df_print: paged
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      comment = "#>", highlight = TRUE,
                      fig.align = "center")
```

# Packages

```{r}
library(tidyverse)
library(rvest)
```

# Exercise 1 {.tabset .tabset-fade .tabset-pills}

## Problem

Go to http://books.toscrape.com/catalogue/page-1.html and scrape the first 
five pages of data on books with regards to their

1. title
2. price
3. star rating
4. availability

Organize your results in a neatly formatted tibble similar to below.

```{r echo=FALSE}
get_books <- function(page) {
  
  base_url <- "http://books.toscrape.com/catalogue/page-"
  url <- str_c(base_url, page, ".html")
  
  books_html <- read_html(url)
  
  prices <- books_html %>% 
    html_nodes(css = ".price_color") %>% 
    html_text()
  
  titles <- books_html %>% 
    html_nodes(css = ".product_pod a") %>% 
    html_attr("title") %>% 
    .[!is.na(.)]

  ratings <- books_html %>% 
    html_nodes(css = ".star-rating") %>% 
    html_attr(name = "class") %>% 
    str_remove(pattern = "star-rating ")
  
  availabilities <- books_html %>% 
    html_nodes(css = ".availability") %>% 
    html_text() %>% 
    str_trim()
  
  books_df <- tibble(
    title     = titles,
    price     = prices,
    rating    = ratings,
    available = availabilities
  )
  
  return(books_df)
}
```

```{r echo=FALSE}
# iterate across pages
pages <- 1:5
books <- map_df(pages, get_books)
books
```

## Solution

```{r}
get_books <- function(page) {
  
  base_url <- "http://books.toscrape.com/catalogue/page-"
  url <- str_c(base_url, page, ".html")
  
  books_html <- read_html(url)
  
  prices <- books_html %>% 
    html_nodes(css = ".price_color") %>% 
    html_text()
  
  titles <- books_html %>% 
    html_nodes(css = ".product_pod a") %>% 
    html_attr("title") %>% 
    .[!is.na(.)]

  ratings <- books_html %>% 
    html_nodes(css = ".star-rating") %>% 
    html_attr(name = "class") %>% 
    str_remove(pattern = "star-rating ")
  
  availabilities <- books_html %>% 
    html_nodes(css = ".availability") %>% 
    html_text() %>% 
    str_trim()
  
  books_df <- tibble(
    title     = titles,
    price     = prices,
    rating    = ratings,
    available = availabilities
  )
  
  return(books_df)
}
```

```{r}
# iterate across pages
pages <- 1:5
books <- map_df(pages, get_books)
books
```

# Exercise 2 {.tabset .tabset-fade .tabset-pills}

## Problem

Scrape the first page of books from each genre in the side bar on the website
http://books.toscrape.com/. Scape the same information as before and include
the genre.

```{r echo=FALSE}
base_url <- "http://books.toscrape.com/"
books_main_html <- read_html("http://books.toscrape.com/index.html")

genre_urls <- books_main_html %>% 
  html_nodes(".nav-list ul a") %>%
  html_attr(name = "href") %>% 
  str_c(base_url, .)

genres <- books_main_html %>% 
  html_nodes(".nav-list ul a") %>%
  html_text() %>% 
  str_trim()
```

Modify our function from Exercise 1 to scrape the books by genre.

```{r echo=FALSE}
get_books_by_genre <- function(url, genre) {
  
  books_html <- read_html(url)
  
  prices <- books_html %>% 
    html_nodes(css = ".price_color") %>% 
    html_text()
  
  titles <- books_html %>% 
    html_nodes(css = ".product_pod a") %>% 
    html_attr("title") %>% 
    .[!is.na(.)]

  ratings <- books_html %>% 
    html_nodes(css = ".star-rating") %>% 
    html_attr(name = "class") %>% 
    str_remove(pattern = "star-rating ")
  
  availabilities <- books_html %>% 
    html_nodes(css = ".availability") %>% 
    html_text() %>% 
    str_trim()
  
  books_df <- tibble(
    title     = titles,
    price     = prices,
    rating    = ratings,
    available = availabilities,
    genre     = genre
  )
  
  return(books_df)
}
```

```{r echo=FALSE}
map2_df(genre_urls, genres, get_books_by_genre)
```

## Solution

First, grab all the genre URLs.

```{r}
base_url <- "http://books.toscrape.com/"
books_main_html <- read_html("http://books.toscrape.com/index.html")

genre_urls <- books_main_html %>% 
  html_nodes(".nav-list ul a") %>%
  html_attr(name = "href") %>% 
  str_c(base_url, .)

genres <- books_main_html %>% 
  html_nodes(".nav-list ul a") %>%
  html_text() %>% 
  str_trim()
```

Modify our function from Exercise 1 to scrape the books by genre.

```{r}
get_books_by_genre <- function(url, genre) {
  
  books_html <- read_html(url)
  
  prices <- books_html %>% 
    html_nodes(css = ".price_color") %>% 
    html_text()
  
  titles <- books_html %>% 
    html_nodes(css = ".product_pod a") %>% 
    html_attr("title") %>% 
    .[!is.na(.)]

  ratings <- books_html %>% 
    html_nodes(css = ".star-rating") %>% 
    html_attr(name = "class") %>% 
    str_remove(pattern = "star-rating ")
  
  availabilities <- books_html %>% 
    html_nodes(css = ".availability") %>% 
    html_text() %>% 
    str_trim()
  
  books_df <- tibble(
    title     = titles,
    price     = prices,
    rating    = ratings,
    available = availabilities,
    genre     = genre
  )
  
  return(books_df)
}
```

```{r}
map2_df(genre_urls, genres, get_books_by_genre)
```


# Exercise 3 {.tabset .tabset-fade .tabset-pills}

## Problem

HTML tags are composed of three things: an opening tag, content and ending tag.
They each have different properties. Identify what the following tags are used
for. Only the opening tag is shown.

|    Tag    | Description |
|:---------:|-------------|
|   `<b>`   |             |
|   `<i>`   |             |
|   `<h3>`  |             |
| `<table>` |             |
|   `<tr>`  |             |
|   `<th>`  |             |
|   `<td>`  |             |
|  `<img>`  |             |
|   `<p>`   |             |

## Solution

HTML tags are composed of three things: an opening tag, content and ending tag.
They each have different properties. Identify what the following tags are used
for. Only the opening tag is shown.

|    Tag    | Description         |
|:---------:|---------------------|
|   `<b>`   | bold                |
|   `<i>`   | italics             |
|   `<h3>`  | level 3 header      |
| `<table>` | table               |
|   `<tr>`  | row in a table      |
|   `<th>`  | header in a table   |
|   `<td>`  | cell in a table     |
|  `<img>`  | image               |
|   `<p>`   | paragraph           |