---
title: "Exercises: Databases and SQL"
author: "Shawn Santo"
date: ""
output: 
  html_document:
    css: "exercises.css"
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: false
    df_print: paged
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      comment = "#>", highlight = TRUE,
                      fig.align = "center")
```

# Packages

```{r}
library(DBI)
library(dbplyr)
library(dplyr)
library(nycflights13)
library(Lahman)
```

# Data

Create an in-memory database and add `airports` from `nycflights13` to the
database.

```{r}
con <- DBI::dbConnect(RSQLite::SQLite(), dbname = ":memory:")
copy_to(con, df = nycflights13::airports, "airports")
```

```{r}
airports_db <- tbl(con, "airports")
```

Also, add some baseball data from the `Lahman` package to our 
in-memory database.

```{r}
dbWriteTable(con, "batting", Batting)
dbWriteTable(con, "pitching", Pitching)
dbWriteTable(con, "teams", Teams)
```

# Exercise 1 {.tabset .tabset-fade .tabset-pills}

## Problem

What are the corresponding SQL verbs based on the `dplyr` structure below?

```{r}
airport_car <- airports_db %>% 
  filter(lat >= 33.7666, lat <= 36.588, lon >= -84.3201, lon <= -75.4129) %>% 
  arrange(desc(alt)) %>% 
  select(name, alt)
```

## Solution

```{r}
airport_car %>% 
  show_query()
```

# Exercise 2 {.tabset .tabset-fade .tabset-pills}

## Problem

Add `Salaries` from package `Lahman` as a table to your in-memory database.

## Solution

```{r}
dbWriteTable(con, "salaries", Salaries)
```

# Exercise 3 {.tabset .tabset-fade .tabset-pills}

## Problem

Compute the team salaries for each team in 2016 and display the 5 teams with
the highest payroll. Which team had the lowest payroll in that year?

## Solution

```{r}
dbGetQuery(con, paste("SELECT teamID, SUM(salary) as payroll",
                      "FROM salaries",
                      "WHERE yearID = 2016",
                      "GROUP BY teamID",
                      "ORDER BY payroll DESC",
                      "LIMIT 5"))
```

# Exercise 4 {.tabset .tabset-fade .tabset-pills}

## Problem

Who were the top 10 teams according to win percentage since 1990? *Hint*:
https://www.w3schools.com/sql/func_sqlserver_cast.asp
   
## Solution

```{r}
dbGetQuery(con, paste("SELECT yearID, teamID, W, L, CAST(W as FLOAT) / G as wpct", 
                      "FROM teams",
                      "WHERE yearID >= 1990",
                      "GROUP BY yearID, teamID",
                      "ORDER BY wpct DESC",
                      "LIMIT 10"))
```

# Exercise 5 {.tabset .tabset-fade .tabset-pills}

## Problem

How would you combine the batting and salaries tables to match up the players
and years? Take a look at `?dplyr::join`. Try to combine the R data frame 
objects `Batting` and `Salaries`.

## Solution

```{r}
left_join(Batting, Salaries, by = c("playerID", "yearID"))
```

