---
title: "Exercises: Memory and I/O"
author: "Shawn Santo"
date: ""
output: 
  html_document:
    css: "exercises.css"
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: false
    df_print: paged
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      comment = "#>", highlight = TRUE,
                      fig.align = "center")
```

# Packages

```{r}
library(lobstr)
library(tidyverse)
```

# Exercise 1 {.tabset .tabset-fade .tabset-pills}

## Problem

Can you diagnose what is going on below?

```{r}
x <- 1:10
y <- x

tracemem(x)
c(obj_addr(x), obj_addr(y))

y[1] <- 3
```

## Solution

The question is, why are two copies being made? The vector `x` is of type
integer. However, when we do subassignment and change the first component of
`y` to be 3 (of type double) two copies are made. One for the modification of 
the component, the other for the atomic vector type change.

```{r}
x <- 1:10
y <- x

tracemem(x)
c(obj_addr(x), obj_addr(y))

y[1] <- 3L # type integer
```


# Exercise 2 {.tabset .tabset-fade .tabset-pills}

## Problem

Starting from 0 we can see that

```{r}
lobstr::obj_size(integer(0))
lobstr::obj_size(numeric(0))
```

are both 48 bytes. Run the code below and see if you can deduce how 
R handles these numeric data in memory?

```{r eval=FALSE}
diff(sapply(0:100, function(x) obj_size(integer(x))))
```

```{r eval=FALSE}
c(obj_size(integer(20)), obj_size(integer(22)))
```

```{r eval=FALSE}
diff(sapply(0:100, function(x) obj_size(numeric(x))))
```

```{r eval=FALSE}
c(obj_size(numeric(10)), obj_size(numeric(14)))
```

## Solution

R allocates memory to vectors in chunks. An integer vector of length one is
allocated 56 bytes, 8 more than a null integer vector. Since an integer
component only requires 4 bytes of memory, an integer vector of length two is
also only 56 bytes. R does not need any more memory. Hence, we see that
`obj_size(integer(1))` and `obj_size(integer(2))` are the same. The `diff()`
function calls give you an idea as to how memory is allocated in chunks.
