---
title: "Exercises: Data structures and subsetting"
author: "Shawn Santo"
date: ""
output: 
  html_document:
    css: "exercises.css"
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: false
    df_print: paged
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      comment = "#>", highlight = TRUE,
                      fig.align = "center")
```

# Exercise 1 {.tabset .tabset-fade .tabset-pills}

## Problem

Create a 3 x 3 x 2 array using the `dim` attribute with the vector below.

```{r}
x <- c(5, 1, 5, 5, 1, 1, 5, 3, 2, 3, 2, 6, 4, 4, 1, 2, 1, 3)
```

Try to create the same array using function `array()`. What do you notice about
how the array object is populated?

## Solution

```{r}
x <- c(5, 1, 5, 5, 1, 1, 5, 3, 2, 3, 2, 6, 4, 4, 1, 2, 1, 3)

# set dim attribute
attr(x = x, which = "dim") <- c(3, 3, 2)
x
attributes(x)
```

```{r}
array(x, dim = c(3, 3, 2))
```

The array is populated using column-major ordering.

| vector index | row | column | matrix |
|--------------|-----|--------|--------|
| 1            | 1   | 1      | 1      |
| 2            | 2   | 1      | 1      |
| 3            | 3   | 1      | 1      |
| 4            | 1   | 2      | 1      |
| 5            | 2   | 2      | 1      |
| 6            | 3   | 2      | 1      |
| 7            | 1   | 3      | 1      |
| 8            | 2   | 3      | 1      |
| 9            | 3   | 3      | 1      |
| 10           | 1   | 1      | 2      |
| 11           | 2   | 1      | 2      |
| 12           | 3   | 1      | 2      |
| 13           | 1   | 2      | 2      |
| 14           | 2   | 2      | 2      |
| 15           | 3   | 2      | 2      |
| 16           | 1   | 3      | 2      |
| 17           | 2   | 3      | 2      |
| 18           | 3   | 3      | 2      |

# Exercise 2 {.tabset .tabset-fade .tabset-pills}

## Problem

Create a factor vector based on the vector of airport codes below. Try to do
it without using function `factor()`.

```{r}
airports <- c("RDU", "ABE", "DTW", "GRR", "RDU", "GRR", "GNV",
             "JFK", "JFK", "SFO", "DTW")
```

Assume all the possible levels are 

```{r eval=FALSE}
c("RDU", "ABE", "DTW", "GRR", "GNV", "JFK", "SFO")
```

*Hint*: Think about what type of object factors are built on.

<br/>

What if the possible levels are

```{r eval=FALSE}
c("RDU", "ABE", "DTW", "GRR", "GNV", "JFK", "SFO", "GSO", "ORD", "PHL")
```

## Solution

Factors are built on atomic integer vectors. First we'll create an integer
vector that serves as a mapping to the unique airport codes.

```{r}
z <- as.integer(c(1,2,3,4,1,4,5,6,6,7,3))

# set levels
attr(x = z, which = "levels") <- c("RDU", "ABE", "DTW", 
                                   "GRR", "GNV", "JFK", "SFO")

# set class to be factor
attr(x = z, which = "class") <- "factor"
z
attributes(z)
```

If the levels change, then we need to adjust our integer vector to have 9
integer values.

# Exercise 3 {.tabset .tabset-fade .tabset-pills}

## Problem

Consider the vectors `x` and `y` below.

```{r}
x <- letters[1:5]
y <- list(i = 1:5, j = -3:3, k = rep(0, 4))
```

What is difference between subsetting with `[` and `[[` using integers? Try
various indices.

## Solution

Let's look at atomic vectors first.

```{r error=TRUE}
x[[4]]
x[4]

x[[1:3]]
x[1:3]
```

Subsetting with a vector of length one appears to be the same using
`[` versus `[[`. We cannot use `[[` and subset with a vector of length greater
than one.

Doing the same for `y` yields

```{r error=TRUE}
y[[2]]
y[2]

y[[1:2]]
y[1:2]

y[[1:3]]
y[1:3]
```

We see similar results for the generic vector `y`. We still should not use
`[[` when subsetting with a vector of length greater than one. The result of
`y[[1:2]]` seems to work, but it is not doing what you think. That code 
is equivalent to `y[[1]][2]`.

# Exercise 4 {.tabset .tabset-fade .tabset-pills}

## Problem

Consider the atomic vector `x`.

```{r}
set.seed(73961)
x <- sample(1:100, size = 100, replace = TRUE)
x
```

Use subsetting to

1. select every third value from `x` beginning at position 6;

2. remove all values with an odd index;

3. remove all numbers divisible by 3 or 5 and replace them with 0.

## Solution

Recall that `x` is

```{r echo=FALSE}
x
```


#### Part 1

```{r}
x[seq(6, length(x), by = 3)]
```

#### Part 2

```{r}
x[-seq(1, length(x), 2)]
```

#### Part 3

```{r}
x[!(x %% 3) | !(x %% 5)] <- 0
x
```

# Exercise 5 {.tabset .tabset-fade .tabset-pills}

## Problem

Consider the list given below.

```{r}
x <- jsonlite::fromJSON('{
  "id" : "8671703e-aab7-47a8-818e-a83a91278658",
  "index" : 1,
  "period" : 1,
  "timestamp" : "00:00:00.000",
  "minute" : 0,
  "second" : 0,
  "type" : {
    "id" : 35,
    "name" : "Starting XI"
  },
  "possession" : 1,
  "possession_team" : {
    "id" : 746,
    "name" : "Manchester City WFC"
  }}')

str(x)
```

1. Subset `x` to obtain the `period`, `minute`, and `second` as a list.

2. Subset `x` to obtain the `possession_team` `name` as a character vector.

3. Subset `x` to obtain a list with `type` as the first component of the list.

## Solution

#### Part 1

```{r}
x[c("period", "minute", "second")]
```

#### Part 2

```{r}
x$possession_team$name
```

#### Part 3

```{r}
x["type"]
```

