---
title: "Exercises: Spatial data visualization with sf"
author: "Shawn Santo"
date: ""
output: 
  html_document:
    css: "exercises.css"
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: false
    df_print: paged
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      comment = "#>", highlight = TRUE,
                      fig.align = "center")
```

# Packages

```{r}
library(sf)
library(tidyverse)
```

# Data

## NC counties

```{r}
nc <- st_read(system.file("shape/nc.shp", package = "sf"), quiet = TRUE)
```

# Exercise 1 {.tabset .tabset-fade .tabset-pills}

## Problem

Use `ggplot` to create a choropleth map for the proportion of sudden infant 
deaths, for the period of July 1, 1974 to June 30, 1979.

```{r echo=FALSE}
nc %>% 
  select(BIR74, SID74) %>% 
  mutate(SID74_prop = SID74 / (BIR74 + SID74)) %>% 
  ggplot() +
  geom_sf(aes(fill = SID74_prop)) +
  scale_fill_gradient(low = "#fff7f3", high = "#49006a") +
  labs(title = "July 1, 1974 to June 30, 1979", 
       fill  = "",
       subtitle = "Proportion of SID by county") +
  theme_void() +
  theme(plot.margin = margin(0, 1, 0, 1, "cm"))
```


## Solution

```{r}
nc %>% 
  select(BIR74, SID74) %>% 
  mutate(SID74_prop = SID74 / (BIR74 + SID74)) %>% 
  ggplot() +
  geom_sf(aes(fill = SID74_prop)) +
  scale_fill_gradient(low = "#fff7f3", high = "#49006a") +
  labs(title = "July 1, 1974 to June 30, 1979", 
       fill  = "",
       subtitle = "Proportion of SID by county") +
  theme_void() +
  theme(plot.margin = margin(0, 1, 0, 1, "cm"))
```

# Exercise 2 {.tabset .tabset-fade .tabset-pills}

## Problem

Recreate John Snow's cholera map showing the cholera death locations and the
water pump locations.

Download the data (you should have a directory named `data/`):

```{r eval=FALSE}
download.file("http://rtwilson.com/downloads/SnowGIS_SHP.zip",
              destfile = "data/john_snow.zip")
unzip("data/john_snow.zip", exdir = "data/")
```

Read in the data:

```{r}
cholera <- st_read("data/SnowGIS_SHP/Cholera_Deaths.shp", quiet = TRUE)
pumps <- st_read("data/SnowGIS_SHP/Pumps.shp", quiet = TRUE)
```

Before you start, check that both `sf` objects have the same CRS.

<br/>

```{r echo=FALSE}
js_raster <- raster::raster("data/SnowGIS_SHP/OSMap_Grayscale.tif")
```

```{r echo=FALSE, fig.width=6, fig.height=6}
RStoolbox::ggR(js_raster) +
  geom_sf(data = cholera, aes(size = Count), color = "red", alpha = .5) + 
  geom_sf(data = pumps, color = "blue", size = 3) + 
  labs(size = "Deaths", title = "London Cholera Outbreak",
       caption = "Water pumps colored in blue") +
  theme_void() +
  theme(legend.position  = "bottom",
        plot.title       = element_text(size   = 28, 
                                        family = "Impact"),
        legend.text      = element_text(size   = 14, 
                                        family = "Comic Sans MS"),
        legend.title     = element_text(size   = 16, 
                                        family = "Comic Sans MS"),
        plot.caption     = element_text(size   = 10, 
                                        family = "Comic Sans MS"))
```

## Solution

Here is a way to overlay the points on the raster object.

```{r eval=FALSE}
js_raster <- raster::raster("data/SnowGIS_SHP/OSMap_Grayscale.tif")
```

```{r eval=FALSE}
RStoolbox::ggR(js_raster) +
  geom_sf(data = cholera, aes(size = Count), color = "red", alpha = .5) + 
  geom_sf(data = pumps, color = "blue", size = 3) + 
  labs(size = "Deaths", title = "London Cholera Outbreak",
       caption = "Water pumps colored in blue") +
  theme_void() +
  theme(legend.position  = "bottom",
        plot.title       = element_text(size   = 28, 
                                        family = "Impact"),
        legend.text      = element_text(size   = 14, 
                                        family = "Comic Sans MS"),
        legend.title     = element_text(size   = 16, 
                                        family = "Comic Sans MS"),
        plot.caption     = element_text(size   = 10, 
                                        family = "Comic Sans MS"))
```

Leaflet is a challenge here because CRS are not aligned. R's `plot()` functions
should work too.