---
title: "Exercises: Data manipulation: dplyr"
author: "Shawn Santo"
date: ""
output: 
  html_document:
    css: "exercises.css"
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: false
    df_print: paged
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      comment = "#>", highlight = TRUE,
                      fig.align = "center")
```

# Packages

```{r}
library(tidyverse)
library(janitor)
```

# Data

Parcel boundaries with address and revenue-related information for properties in 
Wake County, NC. http://data-wake.opendata.arcgis.com/datasets/parcels

```{r}
wake <- read_csv("https://www2.stat.duke.edu/~sms185/data/econ/parcels.csv")

wake <- clean_names(wake)
```

# Exercise 1 {.tabset .tabset-fade .tabset-pills}

## Problem

Which city has the fewest land parcels in the dataset?

## Solution

```{r}
wake %>% 
  count(city_decode) %>% 
  arrange(n) %>% 
  slice(1)
```

# Exercise 2 {.tabset .tabset-fade .tabset-pills}

## Problem

Create a tibble that shows the year a parcel was built and the total
value, where all parcels are located in Apex and are more than one acre in
area. Sort the result in ascending order by year built.

## Solution

```{r}
wake %>% 
  filter(city_decode == "APEX", calc_area > 1) %>% 
  select(year_built, total_value_assd) %>% 
  arrange(year_built)
```

# Exercise 3 {.tabset .tabset-fade .tabset-pills}

## Problem

Choose a subset of five variables and 10 random rows from `wake` and save it 
as an object named `wake_mini`. Experiment renaming variables with
`select()` and `rename()` on `wake_mini`. What is the difference between the
two functions?

## Solution

- Function `select()` only keeps the variables you mention.

- Function `rename()` keeps all variables.

# Exercise 4 {.tabset .tabset-fade .tabset-pills}

## Problem

Compute the mean area for each design style.

## Solution

```{r}
wake %>% 
  group_by(design_style_decode) %>% 
  summarise(mean_area = mean(calc_area))
```

# Exercise 5 {.tabset .tabset-fade .tabset-pills}

## Problem

Compute the median sale price for each year. *Hint:* `lubridate::year()`

## Solution

```{r}
wake %>% 
  filter(!is.na(sale_date)) %>% 
  mutate(year = lubridate::year(sale_date)) %>% 
  group_by(year) %>% 
  summarise(median_sale_price = median(totsalprice, na.rm = TRUE))
```

# Exercise 6 {.tabset .tabset-fade .tabset-pills}

## Problem

Which city with at least 1,000 parcels classified as a "Townhouse"
had the highest proportion of parcels as "Townhouse"?

## Solution

```{r}
wake %>% 
  count(city_decode, design_style_decode) %>% 
  group_by(city_decode) %>% 
  mutate(prop = n / sum(n)) %>% 
  ungroup() %>% 
  filter(design_style_decode == "Townhouse", n >= 1000) %>% 
  arrange(desc(prop)) %>% 
  slice(1)
```

