---
title: "Exercises: Functionals"
author: "Shawn Santo"
date: ""
output: 
  html_document:
    css: "exercises.css"
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: false
    df_print: paged
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      comment = "#>", highlight = TRUE,
                      fig.align = "center")
```

# Packages

```{r}
library(tidyverse)
library(repurrrsive)
library(jsonlite)
library(janitor)
```

# Exercise 1 {.tabset .tabset-fade .tabset-pills}

## Problem

Using `sw_people` in package `repurrrsive`, extract the name of all 
characters using:

- a for loop,

- an apply function.

```{r}
str(sw_people[[1]])
```

*Hint:* The `[` and `[[` are functions.

## Solution

Using a `for` loop:

```{r}
out <- character(length = length(sw_people))

for (i in seq_along(sw_people)) {
  out[i] <- sw_people[[i]]$name
}
```

Using `sapply()`:

```{r}
s_out <- sapply(sw_people, `[[`, "name")
```

# Exercise 2 {.tabset .tabset-fade .tabset-pills}

## Problem

Use `mtcars` and a single map or map variant to

1. get the type of each variable,

2. get the fourth row such that result is a character vector,

3. compute the mean of each variable, and

4. compute the mean and median for each variable such that the result is a
   data frame with the mean values in row 1 and the median values in row 2.
   
## Solution

The solutions to 1, 2, 3, 4 are given below by each line of code, respectively.

```{r}
map_chr(mtcars, typeof)
map_chr(mtcars, 4)
map_dbl(mtcars, mean)
map_df(mtcars, ~ c(mean(.), median(.)))
```

# Exercise 3 {.tabset .tabset-fade .tabset-pills}

## Problem

Use a map function and your `mh_distance()` function from Homework 01
to iterate over both vectors `s` and `w` below.
```{r}
s <- c(26, 50123, 456.12, 8, 0)
w <- c(22, 50000, 451.00, 88, 0)
```

## Solutions

```{r}
mh_distance <- function(x, y) {
  # check for correct modes
  disallow_modes <- c("complex", "raw")
  
  if (mode(x) != mode(y) || any(c(mode(x), mode(y)) %in% disallow_modes)) {
    warning("x and y are not both logical, numeric, or character")
    return(-1)
  }
  
  # check not NA, NaN, Inf, -Inf
  if (any(c(is.na(c(x, y)), is.infinite(c(x, y))))) {
    warning("x or y cannot be NA, NaN, Inf, or -Inf")
    return(-1)
  }
  
  # check not decimal
  # if logical or numeric, convert to character
  if (!is.character(x)) {
    x <- abs(as.numeric(x))
    y <- abs(as.numeric(y))
    
    if (any(c(x, y) %/% 1 != c(x, y))) {
      warning("x or y cannot contain decimal values")
      return(-1)
    }
    
    x <- as.character(x)
    y <- as.character(y)
  }
  
  # check same number of digits or letters
  # x and y are both of type character at this step
  if (nchar(x) != nchar(y)) {
    warning("x and y do not have the same number of digits or letters")
    return(-1)
  }
  
  # parse at each letter or number
  x_split <- unlist(strsplit(x, split = ""))
  y_split <- unlist(strsplit(y, split = ""))
  
  return(sum(x_split != y_split))
}
```

```{r}
s <- c(26, 50123, 456.12, 8, 0)
w <- c(22, 50000, 451.00, 88, 0)

map2_dbl(s, w, ~ suppressWarnings(mh_distance(x = .x, y = .y)))
```

# Exercise 4 {.tabset .tabset-fade .tabset-pills}

Consider the object `senators` below.

```{r message=FALSE, warning=FALSE}
json_file <- "https://www.govtrack.us/api/v2/role?current=true&role_type=senator"
senators <- read_json(json_file)
```

## Problem

Use one of the `map_*()` variants to get

1. the name of each senator as a character vector (preview given below),

```{r echo=FALSE}
map_chr(senators$objects, list("person", "name"))[1:5]
```

2. the name of each senator as a list (preview given below),

```{r echo=FALSE}
map(senators$objects, list("person", "name"))[1:3]
```

3. the description and party of each senator as a data frame 
  (preview given below)
  
```{r echo=FALSE}
map_df(senators$objects, `[`, c("description", "party")) %>% 
  slice(1:4)
```

## Solution

#### Part 1

```{r eval=FALSE}
map_chr(senators$objects, list("person", "name"))
```

#### Part 2

```{r eval=FALSE}
map(senators$objects, list("person", "name"))
```

#### Part 3

```{r eval=FALSE}
map_df(senators$objects, `[`, c("description", "party"))
```
