---
title: "Exercises: Subsetting and S3 OO system"
author: "Shawn Santo"
date: ""
output: 
  html_document:
    css: "exercises.css"
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: false
    df_print: paged
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      comment = "#>", highlight = TRUE,
                      fig.align = "center")
```

# Exercise 1 {.tabset .tabset-fade .tabset-pills}

## Problem

1. Use function `sloop::ftype()` to see which of the following functions are
   S3 generics: `mean`, `summary`, `print`, `sum`, `plot`, `View`, `length`, 
   `[`.

2. Choose 2 of the S3 generics you identified above. How many methods exist for
   each? Use function `sloop::s3_methods_generic()`.
   
3. How many methods exist for classes `factor` and `data.frame`. 
   Use function `sloop::s3_methods_class()`.
   
4. Consider a class called accounting. If a numeric vector has this class,
   function `print()` should print the vector with a $ in front of each number
   and display values up to two decimals. Create a method for this class.
   
## Solution

#### Part 1

```{r}
library(sloop)
```

Checking a couple of the functions:

```{r}
ftype(mean)
ftype(plot)
```

#### Part 2

```{r}
nrow(s3_methods_generic("mean"))
nrow(s3_methods_generic("plot"))
```

#### Part 3

```{r}
nrow(s3_methods_class("factor"))
nrow(s3_methods_class("data.frame"))
```

#### Part 4

```{r}
print.accounting <- function(x) {
  print(paste0("$", format(round(x, digits = 2), nsmall = 2)), quote = FALSE)
}
```

```{r}
x <- 1:5
class(x) <- "accounting"
print(x)
```

```{r}
y <- c(4.292, 134.1133, 50.111)
class(y) <- "accounting"
print(y)
```