---
title: "Exercises: Spatial data wrangling with sf"
author: "Shawn Santo"
date: ""
output: 
  html_document:
    css: "exercises.css"
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: false
    df_print: paged
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      comment = "#>", highlight = TRUE,
                      fig.align = "center")
```

# Packages

```{r}
library(tidyverse)
library(sf)
library(janitor)
library(tidygeocoder)
```

# Data

## Access

## Access data

Create a folder `data/`.

NC PMTW Streams:

```{r eval=FALSE}
download.file(paste0("https://opendata.arcgis.com/datasets/", 
                    "0cc135e6c6244c9e9646b45ee3cb6c1e_0.zip", 
                    "?outSR=%7B%22latestWkid%22%3A3857%2C%22",
                    "wkid%22%3A102100%7D"),
              destfile = "data/pmtw.zip")

unzip("data/pmtw.zip", exdir = "data/")
file.remove("data/pmtw.zip")
```

NC Public Schools:

```{r eval=FALSE}
download.file(paste0("https://opendata.arcgis.com/datasets/", 
                    "dea6ff0e8b4743a0ba361e13a85a4c70_3.zip", 
                    "?outSR=%7B%22latestWkid%22%3A32119%2C%22",
                    "wkid%22%3A32119%7D"),
              destfile = "data/schools.zip")

unzip("data/schools.zip", exdir = "data/")
file.remove("data/schools.zip")
```

## Read

Read in the shapefiles

```{r}
nc_counties <- st_read(system.file("shape/nc.shp", package = "sf"), 
                       quiet = T) %>% 
  clean_names() %>% 
  select(name, area)
```

```{r}
nc_pmtw <- st_read("data/PMTW_Streams_2020.shp", quiet = T) %>%
  clean_names() %>% 
  select(name = displ_name, designation = first_wrc, length = shape_len)
```

```{r}
nc_schools <- st_read("data/Public_Schools.shp", quiet = T) %>% 
  clean_names()
```

## Align CRS

```{r}
nc_counties <- st_transform(nc_counties, crs = st_crs(nc_schools))
nc_pmtw <- st_transform(nc_pmtw, crs = st_crs(nc_schools))
```


```{r}
st_crs(nc_counties)[1]
st_crs(nc_pmtw)[1]
st_crs(nc_schools)[1]
```

# Exercise 1 {.tabset .tabset-fade .tabset-pills}

## Problem

Use `st_intersects()` and `st_contains()` to plot all the public schools in
Durham and the counties that neighbor Durham.

<br/>

```{r echo=FALSE, fig.height=5, fig.width=5}
durham <- nc_counties %>% 
  filter(str_detect(name, "^Durham"))

durham_counties <- durham %>% 
  st_intersects(nc_counties, sparse = FALSE) %>% 
  filter(.data = nc_counties, .)

durham_counties %>% 
  st_contains(nc_schools, sparse = FALSE) %>%
  apply(MARGIN = 2, FUN = any) %>% 
  filter(.data = nc_schools, .) %>% 
  ggplot() +
  geom_sf(color = "blue", alpha = .5, size = 2) +
  geom_sf(data = durham_counties, alpha = 0) +
  labs(title = "Durham Area Public Schools") +
  theme_void()
```

## Solution

```{r fig.height=5, fig.width=5}
durham <- nc_counties %>% 
  filter(str_detect(name, "^Durham"))

durham_counties <- durham %>% 
  st_intersects(nc_counties, sparse = FALSE) %>% 
  filter(.data = nc_counties, .)

durham_counties %>% 
  st_contains(nc_schools, sparse = FALSE) %>%
  apply(MARGIN = 2, FUN = any) %>% 
  filter(.data = nc_schools, .) %>% 
  ggplot() +
  geom_sf(color = "blue", alpha = .5, size = 2) +
  geom_sf(data = durham_counties, alpha = 0) +
  labs(title = "Durham Area Public Schools") +
  theme_void()
```

# Exercise 2 {.tabset .tabset-fade .tabset-pills}

## Problem

Use `st_buffer()` to place a 500 meter buffer around the trout waters in
Ashe county North Carolina.

```{r echo=FALSE, fig.height=7, fig.width=7}
ashe <- nc_counties %>% 
  filter(name == "Ashe")

ashe_trout <- ashe %>% 
  st_contains(nc_pmtw, sparse = FALSE) %>% 
  filter(.data = nc_pmtw, .)

ggplot(ashe) +
  geom_sf(alpha = 0.3) +
  geom_sf(data = st_buffer(ashe_trout, dist = 500), color = "pink", 
          alpha = 0.5) +
  geom_sf(data = ashe_trout, color = "blue") +
  labs(title = "Ashe county PMTW") +
  theme_void()
```

## Solution

```{r fig.height=7, fig.width=7}
ashe <- nc_counties %>% 
  filter(name == "Ashe")

ashe_trout <- ashe %>% 
  st_contains(nc_pmtw, sparse = FALSE) %>% 
  filter(.data = nc_pmtw, .)

ggplot(ashe) +
  geom_sf(alpha = 0.3) +
  geom_sf(data = st_buffer(ashe_trout, dist = 500), color = "pink", 
          alpha = 0.5) +
  geom_sf(data = ashe_trout, color = "blue") +
  labs(title = "Ashe county PMTW") +
  theme_void()
```

# Exercise 3 {.tabset .tabset-fade .tabset-pills}

## Problem

Check the accuracy of the geocoding where `method = census`. Only use a sample
of the schools. You can gauge the accuracy visually or by computing distances
with `st_distance()`. Don't forget to align the CRS!

## Solution

Create a new `sf` object based on the geocode results. We'll take a sample of
10 schools.

```{r}
set.seed(3244)

nc_schools_sample <- nc_schools %>% 
  slice_sample(n = 10)

nc_schools_geocode <- nc_schools_sample %>% 
  st_drop_geometry() %>% 
  select(phys_addr, phys_city, phys_zip) %>% 
  mutate(state = "NC") %>% 
  geocode(street = phys_addr, city = phys_city, state = state, postalcode = phys_zip,
          method = "census") %>%
  filter(!is.na(lat)) %>% 
  st_as_sf(coords = c("long", "lat"), crs = "NAD27") %>% 
  st_transform(crs = st_crs(nc_schools))
```

Visualize how close these are.

```{r}
ggplot(nc_schools_sample) + 
  geom_sf(color = "blue", alpha = .8, size = 5) + 
  geom_sf(data = nc_schools_geocode, color = "red", 
          shape = 21, size = 10) +
  geom_sf(data = nc_counties, alpha = 0) +
  theme_bw()
```



