---
title: "Exercises: Advanced Visualization Techniques"
author: "Shawn Santo"
date: ""
output: 
  html_document:
    css: "exercises.css"
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: false
    df_print: paged
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      comment = "#>", highlight = TRUE,
                      fig.align = "center")
```

# Packages

```{r}
library(tidyverse)
library(patchwork)
```

# Data

**Flint water**

```{r message=FALSE}
url <- str_c("http://www2.stat.duke.edu/~sms185/",
             "data/health/flint.csv")
flint <- read_csv(url)
flint
```

Each row represents a home in Flint, Michigan. Water lead contaminant
value were recorded at three times as represented by `draw1`, `draw2`, and
`draw3`.

# Exercise 1 {.tabset .tabset-fade .tabset-pills}

## Problem

Create a set of visualizations based on tibble object `flint`. Use patchwork
to combine these in a single plot window.

## Solution

```{r fig.align='center', fig.height=9, fig.width=10}
# reshape the data
flint_long <- flint %>% 
  pivot_longer(cols = draw1:draw3, names_to = "draw", values_to = "pb_level")

# create a time plot
p1 <- flint_long %>% 
  filter(zip == 48507, pb_level < 75) %>% 
  ggplot(mapping = aes(x = draw, y = pb_level, group = id)) +
  geom_point() +
  geom_line(color = "grey60") +
  scale_x_discrete(labels = c("Draw 1", "Draw 2", "Draw 3")) +
  labs(y = "Lead level (ppb)", x = "") +
  theme_minimal(base_size = 16)

# create a boxplot
p2 <- flint_long %>% 
  filter(zip == 48507, pb_level < 75) %>% 
  ggplot(mapping = aes(x = draw, y = pb_level)) +
  geom_boxplot() +
  scale_x_discrete(labels = c("Draw 1", "Draw 2", "Draw 3")) +
  labs(y = "", x = "") +
  theme_minimal(base_size = 16)

# create a density plot
p3 <- flint_long %>% 
  filter(zip == 48507, pb_level < 75) %>% 
  ggplot(mapping = aes(x = pb_level, fill = draw)) +
  geom_density(alpha = .3) +
  theme_minimal(base_size = 16) +
  labs(x = "Lead level (ppb)", y = "Density", fill = "Draw time (seconds)") +
  scale_fill_discrete(labels = c("0", "45", "120")) +
  theme(legend.position = "bottom") 

# patchwork to bring everything together
(p1 + p2) / p3
```

# Exercise 2 {.tabset .tabset-fade .tabset-pills}

## Problem

Create a new stat called `stat_outlier_encircle()` that encircles outliers in 
a set of bivariate data. Define an outlier as when both values in a data pair
have a z-score greater than three in absolute value. Check out
https://ggplot2.tidyverse.org/articles/extending-ggplot2.html.

Some examples of `stat_outlier_encircle()` in action are given below.

```{r echo=FALSE}
# create ggproto object
StatOutlier <- ggproto("StatOutlier", Stat,
                       compute_group = function(data, scales) {
                          xvar <- data$x
                          yvar <- data$y
                          
                          stand <- function(x) {
                            (x - mean(x)) / sd(x)
                          }
                          data[abs(stand(xvar)) > 3 & abs(stand(yvar)) > 3, , drop = FALSE]
                       },
                       
                       required_aes = c("x", "y")
                       
                       )

# create stat function
stat_outlier_encircle <- function(mapping = NULL, data = NULL, geom = "point",
                         position = "identity", na.rm = FALSE, show.legend = NA, 
                         inherit.aes = TRUE,
                         shape = 21, size = 5, color = "red",
                         alpha = 1, ...) {
  layer(
    stat = StatOutlier, data = data, mapping = mapping, geom = geom, 
    position = position, show.legend = show.legend, inherit.aes = inherit.aes,
    params = list(color = color, shape = shape, size = size, alpha = alpha, 
                  na.rm = na.rm, ...)
  )
}
```

<br/>

A typical scatter plot with `ggplot()` and `geom_point()`:

```{r}
set.seed(09567899)
data_norm <- tibble(x = c(rnorm(95, 100, 5), rnorm(5, 70, 5)), 
                    y = c(rnorm(95, 100, 5), rnorm(5, 70, 5)))

ggplot(data_norm, mapping = aes(x = x, y = y)) +
  geom_point() +
  theme_minimal()
```

An added layer that marks outliers:

```{r}
ggplot(data_norm, mapping = aes(x = x, y = y)) +
  geom_point() +
  stat_outlier_encircle() +
  theme_minimal()
```

```{r}
ggplot(data_norm, mapping = aes(x = x, y = y)) +
  geom_point() +
  stat_outlier_encircle(color = "purple", size = 14) +
  theme_minimal()
```

```{r}
ggplot(data_norm, mapping = aes(x = x, y = y)) +
  geom_point() +
  stat_outlier_encircle(color = "purple", size = 14, 
                        fill = "lightblue", alpha = .4) +
  theme_minimal()
```

## Solution

```{r}
# create ggproto object
StatOutlier <- ggproto("StatOutlier", Stat,
                       compute_group = function(data, scales) {
                          xvar <- data$x
                          yvar <- data$y
                          
                          stand <- function(x) {
                            (x - mean(x)) / sd(x)
                          }
                          data[abs(stand(xvar)) > 3 & abs(stand(yvar)) > 3, , drop = FALSE]
                       },
                       
                       required_aes = c("x", "y")
                       
                       )

# create stat function
stat_outlier_encircle <- function(mapping = NULL, data = NULL, geom = "point",
                         position = "identity", na.rm = FALSE, show.legend = NA, 
                         inherit.aes = TRUE,
                         shape = 21, size = 5, color = "red",
                         alpha = 1, ...) {
  layer(
    stat = StatOutlier, data = data, mapping = mapping, geom = geom, 
    position = position, show.legend = show.legend, inherit.aes = inherit.aes,
    params = list(color = color, shape = shape, size = size, alpha = alpha, 
                  na.rm = na.rm, ...)
  )
}
```

