---
title: "Exercises: Fundamentals of R"
author: "Shawn Santo"
date: ""
output: 
  html_document:
    css: "exercises.css"
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: false
    df_print: paged
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      comment = "#>", highlight = TRUE,
                      fig.align = "center")
```

# Exercise 1 {.tabset .tabset-fade .tabset-pills}

## Problem

What do each of the following return? Run the code to check your answer.

```{r eval=FALSE}
if (1 == "1") "coercion  works" else "no coercion "

ifelse(5 > c(1, 10, 2), "hello", "olleh")
```

## Solution

```{r}
if (1 == "1") "coercion  works" else "no coercion "
ifelse(5 > c(1, 10, 2), "hello", "olleh")
```

# Exercise 2 {.tabset .tabset-fade .tabset-pills}

## Problem

Consider two vectors, `x` and `y`, each of length one. Write a set of
   conditionals that satisfy the following.
   
- If `x` is positive and `y` is negative or `y` is positive and `x` is
  negative, print "knits".
- If `x` divided by `y` is positive, print "stink".
- Stop execution if `x` or `y` are zero.
    
Test your code with various `x` and `y` values. Where did you
place the stop execution code?

## Solution

```{r}
x <- 4
y <- -10

if (x == 0 | y == 0) {
  stop("One of x or y is 0!")
} else if (x / y > 0) {
  print("stink")
} else {
  print("knits")
}
```

# Exercise 3 {.tabset .tabset-fade .tabset-pills}

## Problem

Consider the vector `x` below.

```{r eval=FALSE}
x <- c(3, 4, 12, 19, 23, 49, 100, 63, 70)
```
    
Write R code that prints the perfect squares in `x`.

## Solution

```{r}
x <- c(3, 4, 12, 19, 23, 49, 100, 63, 70)

for (i in x) {
  if (sqrt(i) %% 1) {
    next
  }
  print(i)
}
```

# Exercise 4 {.tabset .tabset-fade .tabset-pills}

## Problem

Consider `z <- c(-1, .5, 0, .5, 1)`. Write R code that prints
the smallest non-negative integer $k$ satisfying the inequality
$$\lvert cos(k) - z \rvert < 0.001$$
for each component of `z`.

## Solution

```{r}
for (z in c(-1, .5, 0, .5, 1)) {
  k <- 0
  while (abs(cos(k) - z) >= .001) {
    k <- k + 1
  }
  print(k)
}
```

# Exercise 5 {.tabset .tabset-fade .tabset-pills}

## Problem

One way to time your code in R is with `system.time()`. For example,

```{r}
system.time({
  Sys.sleep(3) # suspend code execution for 3 seconds
  runif(n = 1) # generate one uniform(0, 1) random variable
})
```

calculates the elapsed time to be a little over three seconds. Most of the 
elapsed time was due to `Sys.sleep()`. We'll discuss the `user` and `system`
times later in the course.

To quantify the inefficiency of a poorly written `for` loop, time both the 
code blocks below. Experiment with different values of `n`. What do you
observe?

```{r eval=FALSE}
n <- 10
x <- 1
for (i in seq_len(n)) {
  x <- c(x, sqrt(x[i] * i))
}
```

```{r eval=FALSE}
n <- 10
x <- rep(1, n + 1)
for (i in seq_len(n)) {
  x[i + 1] <- sqrt(x[i] * i)
}
```

## Solution

Loop expression 1:

```{r}
system.time({
  n <- 100000
  x <- 1
  for (i in seq_len(n)) {
    x <- c(x, sqrt(x[i] * i))
  }
})
```

Loop expression 2:

```{r}
system.time({
  n <- 100000
  x <- rep(1, n + 1)
  for (i in seq_len(n)) {
    x[i + 1] <- sqrt(x[i] * i)
  }
})
```

```{r echo=FALSE}
library(tidyverse)
time_loops <- function(n) {
  loop_1 <- system.time({
    x <- 1
    for (i in seq_len(n)) {
      x <- c(x, sqrt(x[i] * i))
    }
  })
  
  loop_2 <- system.time({
    x <- rep(1, n + 1)
    for (i in seq_len(n)) {
      x[i + 1] <- sqrt(x[i] * i)
    }
  })
  
  tibble(
    n = n,
    loop_1 = loop_1[3],
    loop_2 = loop_2[3]
  )
  
}
```

```{r echo=FALSE}
map_df(10 ^ (1:5), time_loops) %>% 
  pivot_longer(cols = loop_1:loop_2, names_to = "loop", values_to = "time") %>% 
  ggplot(aes(x = as.factor(n), y = time, color = loop, group = loop)) +
  geom_point(size = 2) +
  geom_line() +
  scale_color_discrete(labels = c("Loop expression 1", "Loop expression 2")) +
  scale_x_discrete(labels = c("10", "100", "1000", "10000", "100000")) +
  labs(x = "Size of n", y = "Elapsed time in seconds", color = "Loop") +
  theme_minimal(base_size = 16)
```


